---
title: "Calculate seed rain by bird species"
author: "Elena Quintero"
date: "`r Sys.Date()`"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
```


```{r, warning=F, message=F}
library(here)
library(tidyverse)
library(tidylog)
library(vegan)
library(magrittr)
library(ggdist)
library(kableExtra)

theme_set(theme_minimal())
```


Load bird's viability:

```{r}
bird.viab <- readRDS(here("data/bird.Pviab_IDs.rds")) %>%
  mutate(bird_sp = str_replace(bird_sp, "_", " ")) %>%
  dplyr::select(-n_bird) %>% 
  arrange(bird_sp)

bird_viab_sum <- bird.viab %>%
  group_by(bird_sp) %>%
  summarise(viab = median(viab)) %>%
  arrange(desc(viab)) 

bird_viab_sum
```


Load mh cover at EP site:

```{r}
mh_cover <- read.csv(here("data/mh_cover.csv"))

mh_cover %<>% dplyr::select(mh = microhab, cover.p = cover.per) %>%
  mutate(mh.cover.m2 = 41000 * cover.p)
  
# El Puntal area (Mancha del Rabicano) where we have worked measures around 41000 sq meters (4.1 hectares)
mh_cover
```


Load posteriors from Bayes models:

1. Number of seeds in rain at EP site:
```{r}
seeds.disp <- readRDS(here("data/Nseed.mh.cover.post.rds"))
glimpse(seeds.disp)
```

2. Prob. that a seed falling in a specific MH comes from a specific bird_sp:

```{r}
p.bird.disp <- readRDS(here("data/Pbird.mh.post.rds"))
glimpse(p.bird.disp)

p.bird.disp %>% 
  group_by(bird_sp) %>% 
  summarise(med = median(Pbird)) %>% 
  arrange(desc(med))
```

Calculate number of seeds dispersed by each bird to each MH:

```{r}
n.mh.bird <- length(unique(seeds.disp$mh))*length(unique(p.bird.disp$bird_sp)) 
# 5 mh * 27 bird spp = 135

n.seed.birds <- p.bird.disp %>% 
  mutate(seeds = case_when(
    mh == "PL" ~ rep(seeds.disp$seeds.m2[seeds.disp$mh == "PL"], times = n.mh.bird),
    mh == "FR" ~ rep(seeds.disp$seeds.m2[seeds.disp$mh == "FR"], times = n.mh.bird),
    mh == "NF" ~ rep(seeds.disp$seeds.m2[seeds.disp$mh == "NF"], times = n.mh.bird),
    mh == "OA" ~ rep(seeds.disp$seeds.m2[seeds.disp$mh == "OA"], times = n.mh.bird),
    mh == "PP" ~ rep(seeds.disp$seeds.m2[seeds.disp$mh == "PP"], times = n.mh.bird)
  )) %>% 
  mutate(seeds.mh.bird = seeds * Pbird)
```

Overview:

```{r, fig.width=10, fig.height=8}
n.seed.birds %>% 
  group_by(bird_sp, mh) %>% 
  summarise(seeds.disp = median(seeds.mh.bird)) %>%
  arrange(desc(seeds.disp))

ggplot(n.seed.birds, aes(y = bird_sp, x = seeds.mh.bird)) + 
  stat_pointinterval() + 
  facet_wrap(~mh, scales = "free")
```

Save dataset:

```{r}
saveRDS(n.seed.birds, here("data/seeds.mh.bird.rds"))
```


Now using this data I select only bird species found in DNA-barcoded feces and sum the total number of seeds contributed by each bird species (sum all MHs), and then multiply by their prob. to disperse a viable seed:

```{r}
glimpse(n.seed.birds)

bc <- read_csv(here("data/seed_rain_by_sampling_point.csv"))

birds_only_cams <- setdiff(unique(n.seed.birds$bird_sp), unique(bc$bird_sp)) 
#remove bird species not present in seed rain, but present in cameras
birds_only_cams

total.bird.mh <- n.seed.birds %>%
  dplyr::select(bird_sp, mh, seeds.mh.bird) %>%
  pivot_wider(names_from = mh, values_from = seeds.mh.bird, values_fn = list) %>%
  tidyr::unnest(cols = c(FR, NF, OA, PL, PP)) %>%
  rowwise() %>%
  mutate(total.seeds.bird = sum(cols = c(FR, NF, OA, PL, PP))) %>%
  arrange(bird_sp) %>%
  cbind(viab = bird.viab$viab) %>% # add p disperse a viab seed 
  mutate(total.viab.seeds.bird = total.seeds.bird * viab) %>% 
  filter(!bird_sp %in% birds_only_cams) #remove species not seen in rain

head(total.bird.mh)
```

Function to calculate PROPORTION of contribution to seed rain:
```{r}
seed.rain.bird.func <- function(dataset, var_col) {
  dataset %>%
  group_by(bird_sp) %>%
  mutate(iter = c(1:1000)) %>%
  group_by(iter) %>%
  mutate(total.seeds = sum({{var_col}})) %>% #get total number of seeds dispersed per iteration
  ungroup() %>%
  mutate(p.disp = {{var_col}} / total.seeds)
}
```

Using any seed:
```{r}
cont.seed.rain <- seed.rain.bird.func(total.bird.mh, total.seeds.bird)

cont.seed.rain.sum <- cont.seed.rain %>% 
  group_by(bird_sp) %>% 
  summarise(median.p.disp = median(p.disp)) %>% 
  arrange(desc(median.p.disp))

sum(cont.seed.rain.sum$median.p.disp) #median proportions closely add up to 1 -good
```

Using only viable seeds:
```{r}
cont.viab.seed.rain <- seed.rain.bird.func(total.bird.mh, total.viab.seeds.bird)

cont.viab.seed.rain.sum <- cont.viab.seed.rain %>% 
  group_by(bird_sp) %>% 
  summarise(median.p.disp.viab = median(p.disp)) %>% 
  arrange(desc(median.p.disp.viab))

sum(cont.viab.seed.rain.sum$median.p.disp.viab) #median proportions closely add up to 1 -good
```

Merge both datasets and clean:
```{r}
bird.seed.rain <- cont.seed.rain %>% 
  dplyr::select(bird_sp, 
         total.seeds.bird,
         total.seeds,
         p.disp, 
         bird.viab = viab,
         total.viab.seeds.bird) %>%
  cbind(total.viab.seeds = cont.viab.seed.rain$total.seeds,
        p.disp.viab = cont.viab.seed.rain$p.disp)

head(bird.seed.rain)

saveRDS(bird.seed.rain, here("data/n.seeds.birds.EP.site.rds"))
```


Contribution to seed dispersal:
```{r}
ggplot(bird.seed.rain, aes(y = bird_sp)) + 
  stat_pointinterval(aes(x=p.disp), color = "blue", point_size = 3, alpha = 0.5) +
  stat_pointinterval(aes(x=p.disp.viab), color = "orange", point_size = 3, alpha = 0.5) +
  labs(subtitle = "Proportion of seed dispersal", x = NULL, y = NULL,
       caption = "Blue is any seed. Orange only VIABLE seeds")
```

```{r}
sessioninfo::session_info()
```

