---
title: "Variation in crop viability data"
author: "Elena Quintero"
output: github_document
---

```{r, message=F}
library(here)
library(tidyverse)
library(assertr)
library(ggplot2)
library(ggpubr)
library(ggrepel)
library(tidylog)
library(RColorBrewer)
```

Load data:

```{r}
viab.data <- read_csv(here("data/viability_collection.csv")) %>%
  mutate(plant_id = as.factor(plant_id))
```


### Viability changes along season in El Puntal year 2019-20:

GLM test:

```{r}
library(glmmTMB)
library(DHARMa)

fit <- glmmTMB(cbind(r_sink+b_sink, r_float+b_float) ~ round + (1|plant_id), 
               family = betabinomial,
               data = subset(viab.data, year=="2019-20" & site=="puntal"))

summary(fit)
simulateResiduals(fit, plot=T)

parameters::parameters(fit, exponentiate = FALSE)
summary(effects::allEffects(fit))
```

```{r}
viab.data %>%
  filter(year == "2019-20" & site == "puntal") %>%
  mutate(term = recode(round, "1" = "1.early", "2" = "2.mid", "3" = "3.late")) %>%
  group_by(term) %>%
  summarise(mean.viab.p = mean(viab.p, na.rm=T), 
            viab.sd = sd(viab.p, na.rm=T),
            viab.se = viab.sd/sqrt(n()),
            n = n()) %>%
  ggplot(aes(x = term, y = mean.viab.p, label = n)) + 
  geom_point() + geom_text(hjust = -0.5, size = 3) +
  #geom_errorbar(aes(ymin=mean.viab.p-viab.sd, ymax=mean.viab.p+viab.sd), width=0.1) +
  geom_errorbar(aes(ymin = mean.viab.p - viab.se, ymax = mean.viab.p + viab.se), width = 0.1) +
  theme_bw() + labs(x = "Collection term in 2019-20", y = "Mean viability", 
                    caption = "Error bars denote 2*SE. 
                    Numbers denote sample size (plants) considered per term")
```


## SUMMARISE DATA PER PLANT PER YEAR

Create a data summary for the 3 round/terms of collection year 2019-20 - either mean for the 3 terms or aggregated terms.

*For the 5 cases where we have wasp.p observations for dry fruits in different collections, I create a "wasp" variable, that preferentially selects the wasp counted over the total fruits (i.e. wasp dry, that also includes the counts in fresh). And a "wasp_ref" variable that indicates the total number of fruits that we need to refer when calculating percentages.*

```{r}
viab.sum <- viab.data %>%
  group_by(site, year, plant_id) %>%
  rename(wasp_fresh=wasp) %>%
  mutate(wasp=ifelse(is.na(wasp_dry), wasp_fresh, wasp_dry)) %>%
  mutate(wasp_ref=ifelse(is.na(wasp_dry), fresh_fruits, total_fruits)) %>%
  summarise(viab.mean=mean(viab.p, na.rm=T), viab.sd=sd(viab.p, na.rm = T),
            total_fruits=sum(fresh_fruits, na.rm=T)+sum(dry, na.rm=T),
            fresh_fruits=sum(fresh_fruits, na.rm=T), 
            dry_fruits=sum(dry, na.rm=T),
            black_v=sum(b_sink, na.rm=T), 
            black_u=sum(b_float, na.rm=T), 
            red_v=sum(r_sink, na.rm=T), 
            red_u=sum(r_float, na.rm=T), 
            abort=sum(abort, na.rm = T), 
            partheno=sum(partheno, na.rm = T),
            wasp=sum(wasp, na.rm = T), wasp_ref=sum(wasp_ref, na.rm = T)) %>%
  mutate(wasp.p=wasp/wasp_ref) %>%
  mutate(viab.agg=(black_v+red_v)/fresh_fruits,
         abort.p=(abort/(abort+partheno))*((1-viab.agg)-wasp.p),
         partheno.p=(partheno/(abort+partheno))*((1-viab.agg)-wasp.p)) %>%
  mutate(black_v_p=black_v/fresh_fruits, 
         black_u_p=black_u/fresh_fruits, 
         red_v_p=red_v/fresh_fruits, 
         red_u_p=red_u/fresh_fruits)
```

Difference between using aggregated viability or mean viability:

Mean viability summary:

```{r}
summary(viab.sum$viab.mean)
sd(viab.sum$viab.mean, na.rm = T)
```

Aggregated viability summary:

```{r}
summary(viab.sum$viab.agg)
sd(viab.sum$viab.agg, na.rm = T)
```

Comparison of estimated viability using mean and aggregated data:

```{r}
ggplot(filter(viab.sum, year=="2019-20" & site=="puntal"), 
       aes(x=viab.mean, y=viab.agg, label=plant_id)) + 
  geom_point() + theme_bw() +  geom_text_repel(size=3) +
  scale_x_continuous(limits = c(0,1)) + scale_y_continuous(limits = c(0,1)) + 
  geom_abline(slope=1) +
  stat_cor()
```

Given there is no huge differences between mean and aggregated viability **it will be more convenient using pooled data, so there is stronger sample power**

```{r}
viab.final <- viab.sum %>% 
  dplyr::select(-c(viab.mean, viab.sd)) %>% 
  rename(viab.p=viab.agg)
```

Clean dataset:
```{r}
viab.for.analysis <- viab.final %>%
  filter(year == "2019-20") %>%
  mutate(black = black_v+black_u, 
         red = red_v+red_u,
         viab = black_v+red_v, 
         non.viab = black_u+red_u) %>%
  rename(id_plant = plant_id,
         total.fruits = total_fruits)

write.csv(viab.for.analysis, here("data/crop_viab.csv"))
```