---
title: "Overview of main results for MS"
author: "Elena Quintero"
date: "`r Sys.Date()`"
output: github_document
---

```{r, message=F}
library(here)
library(tidyverse)
library(tidylog)
library(magrittr)
library(bayestestR)
```



## Plant's viability

```{r}
viab.sum <- read.csv(here("data/crop_viab.csv")) %>% 
  mutate(id_plant = as.factor(id_plant)) 

summary(viab.sum$viab.p)
sd(viab.sum$viab.p)
```

```{r}
viab.sum  %>%
  group_by(site) %>%
  summarise_at(vars(viab.p, abort.p, wasp.p, partheno.p), c(mean, sd))
```


## Consumption

```{r, message=F}
propagules <- readRDS(here("data/propagules_by_plant_and_bird.rds"))
```

Prop. of fruits consumed by main species:
```{r}
sum.fru.cons.bird <- sum.fru.cons %>%
  group_by(bird_sp) %>%
  summarise(fru.cons = sum(fru.cons)) %>%
  arrange(desc(fru.cons))

sum.fru.cons.bird

# First 3 species % of onsumption:
sum(sum.fru.cons.bird$fru.cons[c(1:3)])/sum(sum.fru.cons.bird$fru.cons)
```

Prob. escape avian predation:
```{r}
propagules %>%
  filter(bird_sp == "Chloris_chloris") %>%
  ungroup() %>%
  summarise(median = median(Prob.escape.bird.pred),
            ci = ci(Prob.escape.bird.pred, method = "ETI", ci = 0.8))
```


## Dispersed seed viability

Observed data:
```{r}
viab_seed <- read_csv(here("data/dispersed_seed_viability.csv"))

#Species contributing to seed rain:
viab_seed %>% 
  rowwise() %>%
  mutate(seeds_disp = sum(sink + float)) %>%
  filter(!(sink + float) == 0) %>% 
  distinct(bird_sp) %>%
  print(n = "inf")

# Prob. disp seed being viable
sum(viab_seed$sink) / (sum(viab_seed$sink) + sum(viab_seed$float))

#Number of seeds tested:
sum(viab_seed$sink) + sum(viab_seed$float)
```

Using modelled data - mean seed viability for bird species by migration strategy:
```{r}
source(here("code/bird_migration.R"))

bird.viab <- readRDS(here("data/bird.Pviab_IDs.rds")) %>%
  mutate(bird_type = ifelse(bird_sp %in% resident, "resident", 
                     ifelse(bird_sp %in% winter_migrant, "winter_migrant", "summer_migrant")))

bird.viab %>%
  group_by(bird_type) %>%
  summarise(median = median(viab),
            ci_eti = ci(viab, ci = 0.80))
```

Using modelled data - overall viability of dispersed seeds:
```{r}
general.disp <- propagules %>%
  filter(!is.na(fruits.consumed.sp)) %>%
  mutate(seed.disp = fruits.consumed.sp * Prob.escape.bird.pred,
         viab.seed.disp = fru.cons.viab * Prob.escape.bird.pred) %>%
  group_by(id_plant, bird_sp) %>%
  summarise(seed.disp = median(seed.disp),
            viab.seed.disp = median(viab.seed.disp)) 

sum(general.disp$viab.seed.disp)/sum(general.disp$seed.disp)
```


## MH spatial seed deposition

```{r}
hills <- read_csv(here("data/seed_trap_diversity_hills.csv")) %>% dplyr::select(-q3)

hills
```


## Seed rain density by microhabitats:

Observed data:
```{r}
mh_sampling_effort <- data.frame("PL" = 9.68, 
                                 "FR" = 2.5245, 
                                 "PP" = 2.5245, 
                                 "NF" = 2.5245, 
                                 "OA" = 227.18) %>%
  pivot_longer(cols = everything(), names_to = "mh", values_to = "mh_sampling_effort")

#Using original dataset (not modelled):
sampling_station <- read_csv(here("data/seed_rain_by_sampling_point.csv")) %>%
  filter(bird_sp != "Alectoris rufa") %>% #remove partridge not considered in this study - not seen feeding at p.lentiscus focal plants
  left_join(mh_sampling_effort)

sampling_station %>% 
  group_by(mh) %>%
  summarise(seeds = sum(n_seeds),
            s_eff = first(mh_sampling_effort)) %>% 
  mutate(seeds_per_m2_sampled = seeds /s_eff) %>%
  print(digits = 4)
```

Using dataset from model output:
```{r}
seed.rain <- readRDS(here("data/Nseed.mh.cover.post.rds"))
mh_cover <- read.csv(here("data/mh_cover.csv")) %>% rename(mh = microhab)

seed.rain %<>% left_join(mh_cover) %>%
  mutate(seeds_per_m2 = seeds.m2/cover.m2) 

seed.rain %>% 
  group_by(mh) %>% 
  summarise(median = median(seeds_per_m2),
            ci_80 = ci(seeds_per_m2, ci = 0.80))
```

Number of species contributing to each MH:
```{r}
sampling_station %>% 
  group_by(mh, bird_sp) %>% 
  count() %>% 
  group_by(mh) %>% 
  summarise(birds = n_distinct(bird_sp)) %>%
  mutate(per_sp_detected = birds/20)
  
unique(sampling_station$bird_sp) # total of 20 species contributing with seeds
```

Correlation between seed rain density and bird diversity contribution per sampling point:
```{r}
mh_sampling_point <- sampling_station %>%
  group_by(mh, mh_id) %>%
  filter(mh != "OA") %>%
  summarise(n_seed_m2 = sum(n_seed_m2),
            bird_diversity = n_distinct(bird_sp))

cor.test(mh_sampling_point$n_seed_m2, mh_sampling_point$bird_diversity, conf.level = 0.8)
```


## MH post-dispersal 

```{r}
mh.quality <- readRDS(here("data/mh.quality.rds"))

mh.quality.long <- mh.quality %>% 
  pivot_longer(cols = everything(), names_to = "stage", values_to = "Prob") %>%
  mutate(mh = sub(".*_", "", stage), 
         stage = sub("_.*", "", stage)) 
```

General probability of transition for each stage:
```{r}
mh.quality.long %>%
  group_by(stage) %>%
  summarise(median = median(Prob),
            ci_eti = ci(Prob,ci = 0.80))
```

Probability of transition for each stage separated by microhabitats:
```{r}
mh.quality.long %>%
  group_by(stage, mh) %>%
  summarise(median = median(Prob),
            ci_eti = ci(Prob, method = "ETI", ci = 0.80))
```

Overall probability of recruitment (ORP):
```{r}
mh.ORP <- mh.quality.long %>% 
  pivot_wider(names_from = stage, values_from = Prob, values_fn = list) %>%
  unnest(cols = c(Prob.escaping.pred, 
                  Prob.emergence, 
                  Prob.surv.1summer, 
                  Prob.surv.2summer)) %>%
  mutate(ORP = Prob.escaping.pred * 
               Prob.emergence * 
               Prob.surv.1summer * 
               Prob.surv.2summer)

mh.ORP %>%
  summarise(median = median(ORP),
            ci_eti = ci(ORP, ci = 0.80))

mh.ORP %>%
  group_by(mh) %>%
  summarise(median = median(ORP),
            ci_eti = ci(ORP, ci = 0.80))
```



## Focal plant's recruitment

```{r}
plant.rec <- readRDS(here("data/propagules_by_plant_EP.rds"))
```

Total crop:
```{r}
indv_crops <- plant.rec %>% group_by(id_plant) %>% summarise(crop = median(cropsize))
sum(indv_crops$crop)
summary(indv_crops$crop)
```

Total recruits:
```{r}
plant.rec %>%
  group_by(id_plant) %>%
  mutate(iter = c(1:1000)) %>%
  group_by(iter) %>%
  summarise(recruits = sum(recruits)) %>%
  summarise(median = median(recruits),
            ci_eti = ci(recruits, ci = 0.80))
```

Recruits by individual plants:
```{r}
indv_rec <- plant.rec %>% 
  group_by(id_plant) %>% 
  summarise(recruits = median(recruits),
            viab = median(viab)) %>%
  mutate(rank = rank(desc(recruits))) %>%
  ungroup() %>%
  mutate(total_recruits = sum(recruits),
         prop.rec = recruits/total_recruits) %>%
  arrange(desc(rank)) %>%
  mutate(cum_recruits = cumsum(recruits),
         cum_prop = cumsum(prop.rec)) 

indv_rec %>% arrange(desc(recruits))
```

Correlation between crop viability and no. of recruits:
```{r}
cor.test(indv_rec$viab, indv_rec$recruits, conf.level = 0.8)
```

Fruits required to recruit one seedling:
```{r}
plant.rec %>%
  mutate(fruits.to.recruit.1 = cropsize/recruits) %>%
  group_by(id_plant) %>%
  # mutate(iter = c(1:1000)) %>%
  # group_by(iter) %>%
  summarise(fruits.to.recruit.1 = median(fruits.to.recruit.1)) %>%
  summarise(median = median(fruits.to.recruit.1),
            ci_eti = ci(fruits.to.recruit.1, ci = 0.80))
```


## POPULATION recruitment

Seeds dispersed and recruits in whole population:

```{r}
MH_recruits_POP <- readRDS(here("data/propagules_by_MH_general_EP.rds"))
mh_cover

# total seeds
MH_recruits_POP %>%
  group_by(iter) %>%
  summarise(seed.arrival = sum(seed.arrival))%>%
  summarise(median(seed.arrival),
            ci.80 = ci(seed.arrival, ci = 0.80))

# total recruits
MH_recruits_POP %>%
  group_by(iter) %>%
  summarise(recruits = sum(seedling.surv2))%>%
  summarise(median(recruits),
            ci.80 = ci(recruits, ci = 0.80))
```

Total seeds and recruits by MH:
```{r}
# total seeds dispersed by MH
MH_recruits_POP %>%
  group_by(mh) %>%
  summarise(seeds.dispersed = median(seed.arrival),
            ci_eti = ci(seed.arrival, ci = 0.80))

# total recruits by MH
MH_recruits_POP %>%
  group_by(mh) %>%
  summarise(median = median(seedling.surv2),
            ci_eti = ci(seedling.surv2, ci = 0.80))
```

Number of recruits per every 1000000 seeds at each MH:
```{r}
MH_recruits_POP %>%
  mutate(prob.rec = seedling.surv2/seed.arrival,
         prob.rec.1.000.000 = seedling.surv2/seed.arrival*1000000) %>%
  group_by(mh) %>%
  summarise(median = median(prob.rec.1.000.000),
            ci_eti = ci(prob.rec.1.000.000, ci = 0.80))
```

