---
title: "Probability of seedling survival to 2nd summer at each microhabitat"
date: "`r Sys.Date()`"
output: github_document
---


```{r setup, include=FALSE}
set.seed(888)
library(knitr)
knitr::opts_chunk$set(echo = TRUE, out.width = '100%', 
                      cache = FALSE, message = FALSE)
```

```{r }
library(here)
library(readr)
library(dplyr)
library(ggplot2)
suppressPackageStartupMessages(library(tidylog))
library(DHARMa.helpers)
# source(here("R/check_brms.R"))


theme_set(theme_minimal())
```


Load data

```{r}
mhdata <- read_csv(here("data/seedling_emergence.csv"), guess_max = 10000)

unique(mhdata$revision_date)
```

Select only seeds that survived first summer:
```{r}
survivers <- mhdata %>%
  filter(revision_date == as.Date("2019-10-23") & alive == 1 |
         revision_date == as.Date("2020-09-24") & alive == 1) %>%
  distinct(seed_id)

mhdata_surv <- mhdata %>%
  filter(seed_id %in% survivers$seed_id) %>%
  mutate(max_age = death_date - emergence_date)

mhdata_surv %>% group_by(seed_id) %>% 
  summarise(max_age = first(max_age)) %>% 
  arrange(max_age) %>% print(n = "inf")
```

43 seedlings survived the first summer. 
The seedling that lasted the least was 2018-19_PL324_5A and lived for 266 days, while the longest lasing seed was 2019-20_PL335_4A that lived for 841 days (~ 28 months) (last revision date was 2022-09-27)

If we consider *prob. of seedlings to reach 2-year-old* we would only have 2 surviving seeds:
```{r}
mhdata_surv %>% group_by(seed_id) %>% filter(max_age >= 730) %>% 
  summarise(min_date = min(revision_date)) %>% print(n = "inf")
```

But, if considering *prob. of seedlings surviving 2nd summer* (using mid-October census as final date to consider survival) we would have 6 surviving seeds:
```{r}
mhdata_survived <- mhdata_surv %>%
  filter(season == "2018-19" & revision_date == as.Date("2020-10-14") |
         season == "2019-20" & revision_date == as.Date("2021-09-03")) %>% 
  mutate(year = ifelse(season == "2018-19", "2019", "2020")) %>% 
  mutate(alive = as.numeric(alive))  # censored now coded as NA

table(mhdata_survived$season, mhdata_survived$sowing_date)
table(mhdata_survived$season, mhdata_survived$revision_date)
table(mhdata_survived$season, mhdata_survived$germin_station)
table(mhdata_survived$season, mhdata_survived$alive)
```

Of these 43 seedlings, 18 were from 2018-19 and 25 from 2019-20. For the 18 seedlings in 2018-19, only 1 was alive after the 2nd summer, 6 died and 11 were censored (boar destruction). For the 25 seedlings in 2019-20, 20 died and 5 remained alive.

Censored seeds were in the following microhabitats:
```{r}
mhdata_survived %>% 
  filter(is.na(alive)) %>%
  group_by(season, germin_station) %>% 
  summarise(n())
```

Summary of dead/alive by MH:
```{r}
mhdata_survived <- mhdata_survived %>% 
  filter(!is.na(alive)) %>%
  group_by(season, germin_station) %>%
  mutate(emerged_seeds = n())

# Summary
table(mhdata_survived$mh, mhdata_survived$alive, mhdata_survived$season)
```

The model excludes the censored stations, so I remove the censored stations (-11). Now the n is 32 seedlings, and not 43 as initially selected.


## Model seedling survival to 2nd summer

We model, for every seed that survived the first summer, the probability of surviving the second summer, across microhabitats and years (2018-19 & 2019-20).

There are 5 microhabitats (MH), with variable number of surviving seeds per MH.


### Prior Predictive Check


```{r }
library("brms")

surv.formu <- brmsformula(alive ~ mh + year + (1 | germin_station), 
                            family = bernoulli())

get_prior(surv.formu, data = mhdata_survived)

## priors (logit scale)
# plogis(-0.5) = 0.38  (40% survival)
# plogis(-0.8) = 0.31  (30% survival)
# plogis(-1.3) = 0.21  (20% survival)

plot(hist(plogis(rnorm(10000, -0.8, 2)), breaks = 100))
# plot(hist(plogis(rnorm(10000, -0.8, 3)), breaks = 100))
# plot(hist(plogis(rnorm(10000, -0.8, 2.5)), breaks = 100))

plogis(-3.3)
plogis(-5.3)
plogis(0.7)
plogis(2.7)

priors <- c(
  prior(normal(-0.8, 2), class = "Intercept"), 
  prior(normal(0, 2), class = "b"),
  prior(normal(0, 1), class = "sd")
)
```

Using prior of 20% survival - Really unknown, but would expect something like that. In Gomez-Aparicio 2008 J Ecol the prob of seedling survival is between 0-2-0.4. The prob of 2 years old surviving increases to 0.4 - 0.8. However, since here we are measuring survival to 2nd summer, where seeds are


Prior predictive check:
```{r eval=TRUE}
surv.m <- brm(surv.formu,
         data = mhdata_survived,
         prior = priors,
         sample_prior = "only",
         cores = 4,
         iter = 2000,
         refresh = 0)

# stancode(surv.m)
pp_check(surv.m, type = "bars", ndraws = 100) 
summary(surv.m, priors = TRUE)
plot(surv.m, ask = FALSE)
```



### Fit model


```{r }
surv.m <- brm(surv.formu,
         data = mhdata_survived, 
         prior = priors,
         cores = 4,
         iter = 2000,
         refresh = 0)
         
```


Check: 

```{r }
library(DHARMa)

pp_check(surv.m, type = "bars", ndraws = 100)
bayestestR::check_prior(surv.m)

summary(surv.m, priors = TRUE)

plot(surv.m, ask = FALSE)

#stancode(surv.m)

# coef(surv.m)
fixef(surv.m)

modelbased::estimate_expectation(surv.m)

#mcmc_plot(surv.m, pars = "sampling_point")
# launch_shinystan(surv.m)

#DHARMa
surv.check <- DHARMa.helpers::dh_check_brms(surv.m, integer = TRUE)
# surv.check <- check_brms(surv.m, integer = TRUE)
plot(surv.check, form = as.factor(mhdata_survived$mh))
plot(surv.check, form = as.factor(mhdata_survived$year))
# testCategorical(surv.check, seeds.mh$mh)
#testCategorical(surv.check, seeds.mh.M$sampling_point)
testDispersion(surv.check)
# testUniformity(surv.check)

saveRDS(surv.m, here("code/models/surv.2nd.summer.mh.model.rds"))

```



### Predict

```{r }

## Full posterior distributions

surv.2019 <- posterior_linpred(surv.m, 
                    newdata = data.frame(mh = sort(unique(mhdata_survived$mh)),
                                        emerged_seeds = 1,
                                        year = "2019"),
                    re.form = NA,
                    ndraws = 1000) %>% 
  as.data.frame() %>%
  setNames(sort(unique(mhdata_survived$mh))) %>% 
  pivot_longer(cols = everything(), names_to = "mh", values_to = "surv.2019") %>% 
  arrange(mh)


surv.2020 <- posterior_linpred(surv.m, 
                    newdata = data.frame(mh = sort(unique(mhdata_survived$mh)),
                                        emerged_seeds = 1,
                                         year = "2020"),
                               re.form = NA,
                                ndraws = 1000) %>% 
  as.data.frame() %>%
  setNames(sort(unique(mhdata_survived$mh))) %>% 
  pivot_longer(cols = everything(), names_to = "mh", values_to = "surv.2020") %>% 
  arrange(mh)

surv.post <- bind_cols(surv.2019, dplyr::select(surv.2020, -mh)) %>% 
  mutate(surv.2019 = plogis(surv.2019),
         surv.2020 = plogis(surv.2020)) %>% 
  mutate(surv.avg = (surv.2019 + surv.2020) / 2)

saveRDS(surv.post, here("data/surv.2nd.summer.post.rds"))
```


```{r}
surv.post %>% 
  group_by(mh) %>% 
  summarise(median(surv.2019), median(surv.2020), median(surv.avg))

ggplot(surv.post) +
  geom_density(aes(surv.avg), fill = "lightgrey") +
  facet_wrap(~mh, ncol = 1, scales = "free_y") +
  theme(axis.text.y = element_blank()) +
  labs(title = "Average of both years")

ggplot(surv.post) +
  geom_density(aes(surv.2019), fill = "lightgrey") +
  facet_wrap(~mh, ncol = 1, scales = "free_y") +
  theme(axis.text.y = element_blank()) +
  labs(title = "2019")

ggplot(surv.post) +
  geom_density(aes(surv.2020), fill = "lightgrey") +
  facet_wrap(~mh, ncol = 1, scales = "free_y") +
  theme(axis.text.y = element_blank()) +
  labs(title = "2020")
```

```{r}
library(ggdist)
surv.post <- readRDS(here("data/surv.2nd.summer.post.rds"))

surv.post %>% group_by(mh) %>%
  summarise(median = median(surv.avg),
            CI.0.05 = quantile(surv.avg, probs = 0.05),
            CI.0.95 = quantile(surv.avg, probs = 0.95)
            )

ggplot(surv.post, aes(x = surv.avg, y = mh)) + stat_pointinterval()
```


```{r}
sessioninfo::session_info()
```

