---
title: "Frugivore role in *P. lentiscus* seed rain"
author: "Elena Quintero"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
```


```{r, warning=F, message=F}
library(here)
library(tidyverse)
library(tidylog)
library(vegan)
library(magrittr)
library(ggspatial)
library(kableExtra)
library(ggrepel)

theme_set(theme_minimal())
```

Load data for birds:
```{r}
bird_propagules <- readRDS(here("data/propagules_by_bird_EP.rds")) %>%
  mutate(bird_sp = str_replace(bird_sp, "_", " "))
```


# 1. % consumption and dispersal (using individual plant data)

```{r}
bird.total.cons <- bird_propagules |> 
  select(bird_sp, fru.cons) |> 
  group_by(bird_sp) |> 
  mutate(iteration = 1:1000)
```

Calculate total consumption per iteration:
```{r}
total.cons <- bird.total.cons |> 
  group_by(iteration) |> 
  summarise(total.cons = sum(fru.cons))

bird.prop.cons <- bird.total.cons |> 
  mutate(total.cons.iter = total.cons$total.cons) |> 
  mutate(p.cons = fru.cons/total.cons.iter)

fru.cons.bird.sum <- bird.prop.cons |> 
  group_by(bird_sp) |> 
  summarise(median.fru.cons = median(fru.cons, na.rm = T),
            CI.0.10.fru.cons = quantile(fru.cons, probs = 0.10),
            CI.0.90.fru.cons = quantile(fru.cons, probs = 0.90),
            
            median.p.cons = median(p.cons),
            CI.0.10.p.cons = quantile(p.cons, probs = 0.10),
            CI.0.90.p.cons = quantile(p.cons, probs = 0.90)) %>%
  arrange(desc(median.fru.cons))

sum(fru.cons.bird.sum$median.p.cons) # Not very close to 1 due to medians... 
```

# 2. % consumption and dispersal (using seed traps data) - POPULATION LEVEL

We use data for EP site so values are seeds by square meter in EP site
```{r}
seed.rain.EP <- readRDS(here("data/n.seeds.birds.EP.site.rds"))
#glimpse(seed.rain.EP)
```

```{r}
seed.rain.sum <- seed.rain.EP %>%
  group_by(bird_sp) %>%
  summarise(median.seeds.disp = median(total.seeds.bird),
            CI.0.10.seeds.disp = quantile(total.seeds.bird, probs = 0.10),
            CI.0.90.seeds.disp = quantile(total.seeds.bird, probs = 0.90),
            
            median.p.disp = median(p.disp),
            CI.0.10.p.disp = quantile(p.disp, probs = 0.10),
            CI.0.90.p.disp = quantile(p.disp, probs = 0.90),
            
            median.viab.seeds.disp = median(total.viab.seeds.bird),
            CI.0.10.viab.seeds.disp = quantile(total.viab.seeds.bird, probs = 0.10),
            CI.0.90.viab.seeds.disp = quantile(total.viab.seeds.bird, probs = 0.90),
            
            median.p.disp.viab = median(p.disp.viab),
            CI.0.10.p.disp.viab = quantile(p.disp.viab, probs = 0.10),
            CI.0.90.p.disp.viab = quantile(p.disp.viab, probs = 0.90))

seed.rain.sum %>%
  select(bird_sp, starts_with("median"))

sum(seed.rain.sum$median.p.disp.viab) #sum of median proportions (close to 1)
```

# 3. % of Pistacia in seed rain related to all fruit species
```{r}
all_seeds <- read.csv(here("data/seed_rain_all_fruit_sp.csv"))

p.lentisc.diet <- all_seeds %>%
  group_by(bird_sp, seed_sp1) %>%
  summarise(n.seeds = sum(n_seeds)) %>%
  group_by(bird_sp) %>%
  mutate(total.seeds = sum(n.seeds),
         p.seed.in.rain = n.seeds/total.seeds) %>%
  filter(seed_sp1 == "pistacia") %>%
  arrange(desc(p.seed.in.rain)) 

p.lentisc.diet %>%
  print(n="Inf")
```


# 4. Recruitment by birds (using ind plant data)

```{r}
recruitment <- readRDS(here("data/propagules_by_bird_EP.rds"))
#glimpse(recruitment)
```

```{r}
p.recruitment <- recruitment %>%
  select(bird_sp, bird.total.recruits = recruits) %>%
  mutate(bird_sp = str_replace(bird_sp, "_", " ")) %>%
  group_by(bird_sp) %>%
  mutate(iter = c(1:1000)) %>%
  group_by(iter) %>%
  mutate(total.recruits = sum(bird.total.recruits)) %>%
  ungroup() %>%
  mutate(p.recruitment = bird.total.recruits/total.recruits)

recruitment.sum <- p.recruitment %>%
  group_by(bird_sp) %>%
  summarise(median.recruits = median(bird.total.recruits),
            CI.0.10.recruits = quantile(bird.total.recruits, probs = 0.10),
            CI.0.90.recruits = quantile(bird.total.recruits, probs = 0.90),
            
            median.p.recruit = median(p.recruitment),
            CI.0.10.p.recruit = quantile(p.recruitment, probs = 0.10),
            CI.0.90.p.recruit = quantile(p.recruitment, probs = 0.90)) %>%
  arrange(desc(median.p.recruit))

recruitment.sum

sum(recruitment.sum$median.p.recruit) #good close to 1!
```

# 5. % points where at least 1 seed is contributed by the species

```{r}
sampling_station <- read_csv(here("data/seed_rain_by_sampling_point.csv"))

binary.mat.seed.traps <- sampling_station %>% 
  select(bird_sp, mh_id, n_seed_m2) %>%
  pivot_wider(names_from = mh_id, values_from = n_seed_m2) %>%
  filter(!is.na(bird_sp)) %>% 
  replace(is.na(.), 0) %>%
  mutate_if(is.numeric, ~1 * (. != 0)) 

traps.w.seed <- binary.mat.seed.traps %>%
  rowwise() %>%
  mutate(sum = sum(across(where(is.numeric))),
         p.traps.w.seed = sum/(length(unique(sampling_station$mh_id)))) %>%
  select(bird_sp, n_traps_w_seeds = sum, p.traps.w.seed)

traps.w.seed
```


# 7. diversity sampling points

```{r}
hills <- read_csv(here("data/seed_trap_diversity_hills.csv"))
hills
```

# Frugivore table

```{r}
table <- fru.cons.bird.sum %>%
  left_join(seed.rain.sum) %>% 
  left_join(recruitment.sum) %>%
  #select(-starts_with("CI")) %>%
  left_join(p.lentisc.diet[,c(1,5)]) %>%
  left_join(traps.w.seed[-2]) %>%
  left_join(hills) %>%
  arrange(desc(median.fru.cons)) %>%
  filter(!is.na(median.seeds.disp)) #remove bird with no seed rain data

#glimpse(table)
```

```{r}
table.ms <- table %>%
  select(bird_sp, 
         median.p.cons , CI.0.10.p.cons, CI.0.90.p.cons,
         median.p.disp, CI.0.10.p.disp, CI.0.90.p.disp,
         median.p.disp.viab, CI.0.10.p.disp.viab, CI.0.90.p.disp.viab,
         median.p.recruit , CI.0.10.p.recruit, CI.0.90.p.recruit,
         p.traps.w.seed,
         p.pistacia.seed = p.seed.in.rain) %>%
  arrange(desc(median.p.recruit)) %>%
  mutate(p.cons = paste0(signif(median.p.cons, digits = 1),
                        " [", signif(CI.0.10.p.cons, digits = 1), " – ",
                        signif(CI.0.90.p.cons, digits = 1), "]"),
         p.disp = paste0(signif(median.p.disp, digits = 1),
                        " [", signif(CI.0.10.p.disp, digits = 1), " – ",
                        signif(CI.0.90.p.disp, digits = 1), "]"),
         p.disp.viab = paste0(signif(median.p.disp.viab, digits = 1),
                        " [", signif(CI.0.10.p.disp.viab, digits = 1), " – ",
                        signif(CI.0.90.p.disp.viab, digits = 1), "]"),
         p.recruitment = paste0(signif(median.p.recruit, digits = 1),
                        " [", signif(CI.0.10.p.recruit, digits = 1), " – ",
                        signif(CI.0.90.p.recruit, digits = 1), "]"),
         p.traps.w.seed = round(p.traps.w.seed, digits = 3),
         p.pistacia.seed = round(p.pistacia.seed, digits = 3),
         per.traps.w.seed = p.traps.w.seed * 100, 
         per.pistacia.seed = p.pistacia.seed *100
         ) %>%
  select(bird_sp,
         p.cons, p.disp, p.disp.viab, p.recruitment,
         per.traps.w.seed, per.pistacia.seed) %>%
  mutate(bird_sp = recode(bird_sp, "Sylvia melanocephala" = "Curruca melanocephala",
                       "Sylvia undata" = "Curruca undata",
                       "Sylvia communis" = "Curruca communis",
                       "Sylvia hortensis" = "Curruca hortensis",
                       "Sylvia cantillans" = "Curruca iberiae"))

table.ms %>%
  kbl() %>%
  kable_styling()

write_csv(table.ms, here("data/table.frugivores.csv"))
```

